---
title: "final-project"
format: html
editor: visual
---

## EPI 590R Final Project - Naly Yang

Getting the data:

```{r}
#read in packages
install.packages("tidytuesdayR")
install.packages("tidyverse")
install.packages("gtsummary")
#| include: false
library(tidyverse)
library(tidytuesdayR)
library(gtsummary)

#using here::here
centenarians <- read_csv(here::here("data", "centenarians.csv"))

#or read the csv file manually
centenarians <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-30/centenarians.csv')
```

Recoding data to prepare for regression analysis:
```{r}
#gender variable recoding. if gender= female, then gender1= 1. male= 0.
centenarians$gender1 <- ifelse(centenarians$gender %in% c("female"),1,0)
#still_alive recoding. if still_alive= alive, then still_alive1= 1.deceased= 0.
centenarians$still_alive1 <- ifelse(centenarians$still_alive %in% c("alive"),1,0)
```

## {gtsummary} table of descriptive statistics

```{r}
#| label: tbl-descriptive
#| tbl-cap: "This table contains descriptive statistics."
library(gtsummary)
tbl_summary(
	centenarians,
	by = gender,
	include = c(gender, age, place_of_death_or_residence,
							still_alive),
	label = list(
		gender ~ "Gender",
		age ~ "Age",
		place_of_death_or_residence ~ "Place of Death/Residence",
		still_alive ~ "Living Status"
	),
	missing_text = "Missing")
  footnote= "These are lists of the 100 known verified oldest people sorted in descending order by age in years and days. This data comes from Wikipedia List of verified oldest people via frankiethull on Github. Source= https://github.com/rfordatascience/tidytuesday/tree/master/data/2023/2023-05-30"
```

## Fit a regression and present well-formatted results from the regression

Regression:

```{r}
uni_regression <- tbl_uvregression(
  centenarians,
  y = age,
  include = c(
    gender1, still_alive1,
  ),
  method = lm )
```

Table (you are getting an error here)
```{r}
#| error: true
#| label: tbl-regression
#| tbl-cap: "This table contains regression results."
regress_table <- tbl_regression(
	uni_regression,
	intercept = TRUE,
	label = list(
	  age ~ "Age",
		gender1 ~ "Gender",
		still_alive1 ~ "Living Status"
	))
```

Inline text (you can not resolve this until the error above is resolved):
```{r}
#| error: true
inline_text(uni_regression, variable = "age")
```


## Creating a figure

```{r}
#round age data first. round down to nearest integer & create new vector for the roudned age.
rounded_age <- floor(centenarians$age)
#| label: fig-hist
#| fig-cap: "This is a histogram of the rounded ages."
hist(rounded_age)
```

@fig-hist contains a histogram and @tbl-regression contains a table of the regression results.

## Write and use a function

standard deviation:
```{r}
#testing first by making a dataset
x <- c(20,4,67,100,65,20,17,66,83,61,33,54,22,39,70)
#standard dev steps:
#first find the mean: mean(x)
#find each datapoint's distance to the mean: x-mean
#square of each datapoint's distance to the mean: (x-mean)^2
#sum these distances: sum(x-mean)(x)^2
#divide by n-1  data points in the dataset: (/ length(x)-1)
#double check parenthesis
s <- sqrt(sum(
    (x-mean(x))^ 2 /
      (length(x)-1)))
print(s)
#checking
sd(x)
# it works :) now use on dataset
#use originial "age" variable, not the rounded variable you used earlier in the Figure section.
st_dev <-sqrt(sum
        ((centenarians$age-mean(centenarians$age))^2 /
            (length(centenarians$age)-1)))
print(st_dev)
#checking
sd(centenarians$age)
```

